Music Perception: Stimuli profiles

### 2021-02-23 material/expstimuli/stimuli.Rmd
add mid export
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update stimuli profile
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@

# DVs
- IOIs (average tempo)
- Transition points
- CV (tempo variability)
- KOR (articulation)
- KV (dynamics)
- KV-diff (dynamics)

```{r filtering, include = FALSE}
foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time

# read functions
source("./function.R")

@@

# add column namesls
colnames(raw_data) <- c("NoteNr", "TimeStamp", "Pitch", "Velocity", "Key_OnOff", "Device", "Tempo", "SubNr", "BlockNr", "TrialNr", "Skill", "Condition", "Image")

@@

# correct labelling (due to a labelling error of the original study)
# articulation
dt_a <- raw_data[grepl("stim_a", Image)]
# correct wrong labelling
dt_a$Skill <- "articulation"

# dynamics
dt_d <- raw_data[grepl("stim_d", Image)]
# correct wrong labelling
dt_d$Skill <- "dynamics"

# bind data frames for articulation and dynamics
dt_all <- rbind(dt_a, dt_d)

# raw_data without metronome
dt_all <- dt_all[Pitch != 31 & Pitch != 34]

@@

# convert bpm to ms
dt_ioi[Tempo == 120]$Tempo <- 250
dt_ioi[Tempo == 110]$Tempo <- 273
dt_ioi[Tempo == 100]$Tempo <- 300

# normalise IOIs
dt_ioi$normIOI <- dt_ioi$IOI/dt_ioi$Tempo

ioi <- dt_ioi[Interval != 25 & Interval != 26 & Interval != 27 & Interval != 28 & Interval != 32 & Interval != 33 & Interval != 34 & Interval != 35 & Interval != 51 & Interval != 52 & Interval != 56 & Interval != 57 & Interval != 65 & Interval != 66 & Interval != 67 & Interval != 68 & Interval != 69 & Interval != 70 & Interval != 71, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, BlockNr, TrialNr)]
mean_ioi <- mean(ioi$Mean)
sd_ioi <- sd(ioi$Mean)
ioi_include <- ioi[Mean < mean_ioi+2*sd_ioi & Mean > mean_ioi-2*sd_ioi]

@@

```{r selection, include = FALSE}
# create a list for included performances
list_include <- dt_include_updated[, .N, by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]
list_overall <- list_include[, .N, by = .(Condition, Skill)]

# assign random numbers for performances in each category
list_include$RandNr <- 0
for (cond in c("teaching", "performing")){
  for (skill in c("articulation", "dynamics")){
    list_include[Condition == cond & Skill == skill]$RandNr <- sample(nrow(list_include[Condition == cond & Skill == skill]), replace = FALSE)
  }
}
list_include <- list_include[order(RandNr),]
# stimuli for experimental trials
for (i in 1:96){ # 24 stimuli for each category
  subject = list_include$SubNr[i]
  block = list_include$BlockNr[i]
  trial = list_include$TrialNr[i]
  cond = list_include$Condition[i]
  skill = list_include$Skill[i]
  current <- dt_include[SubNr == subject & BlockNr == block & TrialNr == trial]
  firstTimestamp = current$TimeStamp[1]
  current$TimeStamp <- current$TimeStamp - firstTimestamp # start with time 0

@@

@@

# packages
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}

# create a list of data file names
lf_a_t <- list.files("./1613936591-210221/art_teaching", pattern = "txt")
lf_a_p <- list.files("./1613936591-210221/art_performing", pattern = "txt")
lf_d_t <- list.files("./1613936591-210221/dyn_teaching", pattern = "txt")
lf_d_p <- list.files("./1613936591-210221/dyn_performing", pattern = "txt")

# merge all data files into one
data <- data.table()
for (i in 1:length(lf_a_t)){
 data_i <- fread(file.path("./1613936591-210221/art_teaching", lf_a_t[i]), header = T, sep = ",", dec = ".")
 data_i$Instance <- i
 data <- rbind(data, data_i)
}
for (i in 1:length(lf_a_p)){
 data_i <- fread(file.path("./1613936591-210221/art_performing", lf_a_p[i]), header = T, sep = ",", dec = ".")
 data_i$Instance <- i
 data <- rbind(data, data_i)
}
for (i in 1:length(lf_d_t)){
 data_i <- fread(file.path("./1613936591-210221/dyn_teaching", lf_d_t[i]), header = T, sep = ",", dec = ".")
 data_i$Instance <- i
 data <- rbind(data, data_i)
}
for (i in 1:length(lf_d_p)){
 data_i <- fread(file.path("./1613936591-210221/dyn_performing", lf_d_p[i]), header = T, sep = ",", dec = ".")
 data_i$Instance <- i
 data <- rbind(data, data_i)
}

@@

# IOIs
```{r ioi, include = FALSE}
data_onset <- data[Key_OnOff == 1]
data_offset <- data[Key_OnOff == 0]

data_onset$IOI <- diff(c(0, data_onset$TimeStamp))
dt_ioi <- data_onset[NoteNr != 17]
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# convert bpm to ms
dt_ioi[Tempo == 120]$Tempo <- 250
dt_ioi[Tempo == 110]$Tempo <- 273
dt_ioi[Tempo == 100]$Tempo <- 300

# normalise IOIs
dt_ioi$normIOI <- dt_ioi$IOI/dt_ioi$Tempo

@@

ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(Instance, Condition, Skill)]
```

@@

# KOR
```{r kor, include = FALSE}
data_onset$KOT <- NA
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$TimeStamp[row] - data_onset$TimeStamp[row+1] # offset(n) - onset(n+1)
   }
}

ioi_forKOT <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(Instance, Condition, Skill)]

data_onset$KOR <- 0
for (cond in c("teaching", "performing")){
  for (skill in c("articulation", "dynamics")){
    for (i in c(1:24)){
      data_onset[Condition == cond & Skill == skill & Instance == i]$KOR <- data_onset[Condition == cond & Skill == skill & Instance == i]$KOT/ioi_forKOT[Condition == cond & Skill == skill & Instance == i]$Mean
    }
  }
}

dt_kor <- data_onset[NoteNr != 17]
dt_kor$Interval <- rep(1:71, nrow(dt_kor)/71)

# assign Subcomponents
dt_kor$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
   for (note in 1:length(ls_legato[[phrase]])){
     dt_kor[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
   for (note in 1:length(ls_staccato[[phrase]])){
     dt_kor[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
   }
}

# Forte
for (phrase in 1:length(ls_forte)){
   for (note in 1:length(ls_forte[[phrase]])){
     dt_kor[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano)){
   for (note in 1:length(ls_piano[[phrase]])){
     dt_kor[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
   }
}

# assign Subcomponent Change
for (number in change_1){
   dt_kor[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
   dt_kor[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
   dt_kor[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
   dt_kor[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}

kor <- dt_kor[Subcomponent != "NA" & Subcomponent != "LtoS" & Subcomponent != "StoL" & Subcomponent != "FtoP" & Subcomponent != "PtoF", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(Instance, Condition, Skill, Subcomponent)]
```

## Each instance
```{r kor-each, echo = FALSE}
kor
```

## Averaged
```{r kor-averaged, echo = FALSE}
kor[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), .(Condition, Skill, Subcomponent)]
```

## Plot
### Articulation
```{r plot-kor-art, echo = FALSE}
ggboxplot(kor[Skill == "articulation"], x = "Condition", y = "Mean", color = "Condition", facet.by = "Subcomponent", add = "jitter")
```

### Dynamics
```{r plot-kor-dyn, echo = FALSE}
ggboxplot(kor[Skill == "dynamics"], x = "Condition", y = "Mean", color = "Condition", facet.by = "Subcomponent", add = "jitter")
```

@@

vel <- dt_vel[Subcomponent != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(Instance, Condition, Skill, Subcomponent)]
```

@@

vel_diff <- dt_vel_diff[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(Instance, Condition, Skill, Subcomponent)]
```

>>>>

  + 18 stimuli for teaching-articulation
  + 18 stimuli for teaching-dynamics
  + 18 stimuli for performing-articulation
  + 18 stimuli for performing-dynamics

@@

# DVs
- IOIs (average tempo)
- KOT (articulation)
- KV (dynamics)
- KV-diff (dynamics)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
```

# Filtering
```{r filtering, include = FALSE}
# read functions
source("./function.R")

@@

# add column names
colnames(raw_data) <- c("NoteNr", "TimeStamp", "Pitch", "Velocity", "Key_OnOff", "Device", "Tempo", "SubNr", "BlockNr", "TrialNr", "Skill", "Condition", "Image")

@@

# raw_data without metronome
dt_all <- raw_data[Pitch != 31 & Pitch != 34]

@@

ioi <- dt_ioi[Interval != 25 & Interval != 26 & Interval != 27 & Interval != 28 & Interval != 32 & Interval != 33 & Interval != 34 & Interval != 35 & Interval != 51 & Interval != 52 & Interval != 56 & Interval != 57 & Interval != 65 & Interval != 66 & Interval != 67 & Interval != 68 & Interval != 69 & Interval != 70 & Interval != 71, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Tempo)]
mean_ioi <- mean(ioi$Mean)
sd_ioi <- sd(ioi$Mean)
ioi_include <- ioi[Mean < mean_ioi+2*sd_ioi & Mean > mean_ioi-2*sd_ioi]

@@

## Histogram
- Exclude performances outside 2sd
- Gray vertical lines indicate ideal tempi (300ms for 100bpm, 273ms for 110bpm, 250ms for 120bpm)

1. Before exclusion
```{r plot-hist, echo = FALSE}
ioi$Tempo <- factor(ioi$Tempo)
gghistogram(ioi, x = "Mean", rug = TRUE, add_density = TRUE, bins = 50) + geom_vline(xintercept = 	c(250, 273, 300), color = "gray", size = 1) + lims(y = c(0, 60))
gghistogram(ioi, x = "Mean", fill = "Tempo", color = "Tempo", rug = TRUE, add_density = TRUE, bins = 50) + geom_vline(xintercept = 	c(250, 273, 300), color = "gray", size = 1) + lims(y = c(0, 60))
```

2. After exclusion
```{r plot-hist2, echo = FALSE}
ioi_include$Tempo <- factor(ioi_include$Tempo)
gghistogram(ioi_include, x = "Mean", rug = TRUE, add_density = TRUE, bins = 50) + geom_vline(xintercept = 	c(250, 273, 300), color = "gray", size = 1) + lims(y = c(0, 60))
gghistogram(ioi_include, x = "Mean", fill = "Tempo", color = "Tempo", rug = TRUE, add_density = TRUE, bins = 50) + geom_vline(xintercept = 	c(250, 273, 300), color = "gray", size = 1) + lims(y = c(0, 60))
```

```{r selection, eval = FALSE, include = FALSE}
foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time

# create a list for included performances
list_include <- dt_include_updated[, .N, by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]
list_overall <- list_include[, .N, by = .(Condition, Skill)]

# assign random numbers for performances in each category
list_include$RandNr <- 0
for (cond in c("teaching", "performing")){
  for (skill in c("articulation", "dynamics")){
    list_include[Condition == cond & Skill == skill]$RandNr <- sample(nrow(list_include[Condition == cond & Skill == skill]), replace = FALSE)
  }
}
list_include <- list_include[order(RandNr),]

# stimuli for experimental trials
for (i in 1:72){ # 18 stimuli for each category
  subject = list_include$SubNr[i]
  block = list_include$BlockNr[i]
  trial = list_include$TrialNr[i]
  cond = list_include$Condition[i]
  skill = list_include$Skill[i]
  current <- dt_include[SubNr == subject & BlockNr == block & TrialNr == trial]
  firstTimestamp = current$TimeStamp[1]
  current$TimeStamp <- current$TimeStamp - firstTimestamp # start with time 0

@@

# Export as midi files
```{bash, echo = TRUE}
# python
source /Users/atsukotominaga/Pyenv/bin/activate
python mid_export.py
```

@@

# read midi data with tuneR
# create a list of data file names
lf <- list.files("./mid/", pattern = "mid")

# combine data
all_data <- data.table()
for (file in lf){
   current <- readMidi(paste("./mid/", file, sep = ""))
   current$MidFile <- gsub(".mid", "", file)
   all_data <- rbind(all_data, current)
}

data <- all_data[event == "Note On" | event == "Note Off"]
data$NoteNr <- rep(1:144, nrow(data)/144)
data$channel <- NULL
data$parameterMetaSystem <- NULL
data$track <- NULL
data$type <- NULL
data$time2 <- (data$time/960)*1000 # deltatime to time (ms)

# labelling
colnames(data)[c(3,4)] <- c("Pitch", "Velocity") # time event also change
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", MidFile)]$Condition <- "teaching"
data[grepl("_p_", MidFile)]$Condition <- "performing"
data[grepl("a_", MidFile)]$Skill <- "articulation"
data[grepl("d_", MidFile)]$Skill <- "dynamics"

@@

# IOIs
```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

data_onset$IOI <- diff(c(0, data_onset$time2))
dt_ioi <- data_onset[NoteNr != 1]
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

@@

ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(MidFile, Condition, Skill)]
```

@@

# KOT
```{r kot, include = FALSE}
data_onset$KOT <- NA
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$time2[row] - data_onset$time2[row+1] # offset(n) - onset(n+1)
   }
}

dt_kot <- data_onset[NoteNr != 1]
dt_kot$Interval <- rep(1:71, nrow(dt_kot)/71)

# assign Subcomponents
dt_kot$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
   for (note in 1:length(ls_legato[[phrase]])){
     dt_kot[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
   for (note in 1:length(ls_staccato[[phrase]])){
     dt_kot[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
   }
}

# Forte
for (phrase in 1:length(ls_forte)){
   for (note in 1:length(ls_forte[[phrase]])){
     dt_kot[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano)){
   for (note in 1:length(ls_piano[[phrase]])){
     dt_kot[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
   }
}

# assign Subcomponent Change
for (number in change_1){
   dt_kot[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
   dt_kot[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
   dt_kot[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
   dt_kot[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}

kot <- dt_kot[Subcomponent != "NA" & Subcomponent != "LtoS" & Subcomponent != "StoL" & Subcomponent != "FtoP" & Subcomponent != "PtoF", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile, Condition, Skill, Subcomponent)]
```

## Each instance
```{r kot-each, echo = FALSE}
kot
```

## Averaged
```{r kor-averaged, echo = FALSE}
kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), .(Condition, Skill, Subcomponent)]
```

## Plot
### Articulation
```{r plot-kot-art, echo = FALSE}
ggboxplot(kot[Skill == "articulation"], x = "Condition", y = "Mean", color = "Condition", facet.by = "Subcomponent", add = "jitter")
```

### Dynamics
```{r plot-kor-dyn, echo = FALSE}
ggboxplot(kot[Skill == "dynamics"], x = "Condition", y = "Mean", color = "Condition", facet.by = "Subcomponent", add = "jitter")
```

@@

vel <- dt_vel[Subcomponent != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(MidFile, Condition, Skill, Subcomponent)]
```

@@

vel_diff <- dt_vel_diff[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(MidFile, Condition, Skill, Subcomponent)]
```
